---
title: "Simulations Pre-Class Project"
author: "Pat Zhang"
date: "March 12, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Suppose you have a bankroll of $1000 and make bets of $100 on a fair game. By simulating the outcome directly for at most 5000 iterations of the game (or hands), estimate: 

a. the probability that you have "busted" (lost all your money) by the time you have placed your one hundredth bet.
b. the probability that you have busted by the time you have placed your five hundredth bet by simulating the outcome directly. 
c. the mean time you go bust, given that you go bust within the first 5000 hands. 
d. the mean and variance of your bankroll after 100 hands (including busts). 
e. the mean and variance of your bankroll after 500 hands (including busts).
```{r}
# This function simulates the game and winnings for a specified number of turns
# Inputs: number of hand/games, probability of winning, bankroll, bet/wager, if markov, and increment value
# Output: vector of total money and number of hands/games
sim_game <- function(hands, p = 0.5, inc = 0.01, bankroll = 1000, bets = 100, markov = FALSE) {
  total <- bankroll
  prob <- p
  
  # For each hand/game
  for (i in 1:hands) {
    # Add/subtract wager from total
    outcome <- rbinom(1,1,prob)
    total <- total + (2*bets*outcome - bets)
    # If not Markov, going bust is possible 
    # Return 0 and the index that the player went bust
    if (!markov) {
      if (total == 0) {
        return(c(0,i))
      }
    # If Markov, increment probability accordingly
    } else {
      if (outcome == 1 & prob + inc <= 1) {
        prob <- prob + inc
      } else if (outcome == 0) {
        prob <- p
      }
    }
  }
  return(c(total, hands))
}

# This function calculates the probabilities and means asked in the first two questions.
# Inputs: two values for number of hands/games, probability of winning, and number of repetitions
calc_bust <- function(hands, hands2, prob = 0.5, rep = 5000) {
  # Simulates games for part a and calculates probability of bust
  sim_a <- replicate(rep, sim_game(hands, p = prob))
  prob_a <- length(which(sim_a[1,] == 0))/rep
  # Simulates games for part b and calculates probability of bust
  sim_b <- replicate(rep, sim_game(hands2, p = prob))
  prob_b <- length(which(sim_b[1,] == 0))/rep
  # Simulates games for part c and calculates the average number of hands to bust
  sim_c <- replicate(rep, sim_game(rep, p = prob))
  busted <- which(sim_c[1,] == 0)
  mean_bust <- mean(sim_c[2,busted])
  # Calculates mean and variance for the situations in parts a and b
  mv_d <- c(mean(sim_a[2,]), var(sim_a[2,]))
  mv_e <- c(mean(sim_b[2,]), var(sim_b[2,]))
  return(c(prob_a, prob_b, mean_bust, mv_d, mv_e))
}

sol <- calc_bust(100, 500)
sol
```

Repeat the previous problem with betting on black in American roulette, where the probability of winning on any spin is 18/38 for an even payout.
```{r}
sol2 <- calc_bust(100, 500, prob = 18/38)
sol2
```

Suppose you have a game where the probability of winning on your first hand is 48%; each time you win, that probability goes up by one percentage point for the next game (to a maximum of 100%, where it must stay), and each time you lose, it goes back down to 48%. Assume you cannot go bust and that the size of your wager is a constant $100. 

a. Is this a fair game? Simulate one hundred thousand sequential hands to determine the size of your return. Then repeat this simulation 99 more times to get a range of values to calculate the expectation. 
```{r}
markov_sim <- replicate(100, sim_game(10000, p = 0.48, markov = TRUE))
markov_exp <- mean(markov_sim[1,])
markov_exp
```

b. Repeat this process but change the starting probability to a new value within 2% either way. Get the expected return after 100 repetitions. Keep exploring until you have a return value that is as fair as you can make it. Can you do this automatically? 
c. Repeat again, keeping the initial probability at 48%, but this time change the probability increment to a value different from 1%. Get the expected return after 100 repetitions. Keep changing this value until you have a return value that is as fair as you can make it.
```{r}
# This function determines the fairest starting probability or increment value
# Input: start: TRUE - calculate starting p, FALSE - calculate increment value
# Output: vector of value (index 1) and 100 returns/total money remaining (index 2:101)
fair_markov <- function(start = FALSE) {
  incrm <- 0.01
  prob <- 0.48
  # Smallest difference between starting money and return after playing (absolute value)
  smallest_diff <- abs(1000 - markov_exp)
  repeat {
    # If finding increment, increase value each repetition
    if (!start) {
      incrm <- incrm + 0.001
      sim <- replicate(100, sim_game(10000, p = prob, inc = incrm, markov = TRUE))
      diff <- abs(1000-mean(sim[1,]))
      # If new difference is smaller than smallest_diff, update value
      if (diff < smallest_diff) {
        smallest_diff <- diff
      # If new difference is greater, stop repetition and output
      } else {
        return(c(incrm, sim[1,]))
      }
    # If finding starting p, repeat process above except with increasing prob
    } else {
      prob <- prob + 0.001
      sim <- replicate(100, sim_game(10000, p = prob, markov = TRUE))
      diff <- abs(1000-mean(sim[1,]))
      if (diff < smallest_diff) {
        smallest_diff <- diff
      } else {
        return(c(prob, sim[1,]))
      }
    }
  }
}
start_value <- fair_markov(start = TRUE)
inc_value <- fair_markov()

c(start_value[1], mean(start_value[-1]), var(start_value[-1]))
c(inc_value[1], mean(inc_value[-1]), var(inc_value[-1]))
```

For the last two examples in the previous question, you calculated a mean value. Because you saved these final results in a vector, use the bootstrap to estimate the variance of the return in each case for your final answer. Once you have these results, which game has the smaller variance in returns?
```{r}
boot1 <- start_value[-1]
boot2 <- inc_value[-1]
# Sample both vectors with replacement and take the bootstrap variance
var1 <- var(sample(boot1, replace = TRUE))
var2 <- var(sample(boot2, replace = TRUE))
c(var1, var2)
```

