---
title: "Untitled"
author: ""
date: "March 7, 2017"
output: html_document
---


##Jackknife for Gamma Parameters

Recall our friend the method of moments estimator:

```{r, eval=FALSE}
gamma.est <- function(the_data) {
  m <- mean(the_data)
  v <- var(the_data)
  a <- m^2/v
  s <- v/m
  return(c(a=a,s=s))
}
```


###Jackknife for Gamma Parameters Function

Write a function called `gamma.jackknife` that takes argument `a_vector` and returns jackknife standard error estimates on the gamma parameters. 

```{r, eval=FALSE }
gamma.jackknife <- function(a_vector) {
  n <- length(a_vector)
  
  # Matrix of gamma estimates using omitted data
  theta.i <- matrix(nrow =n, ncol = 2)
  for (row in 1:n) {
    theta.i[row,] <- gamma.est(a_vector[-row])
  }
  
  # Calculates variance of estimates using omitted data
  variance <- apply(theta.i, 2, var)
  
  # Jackknife estimatation for variance and standard error
  jackknife.variance <- (n-1)^2/n * variance
  jackknife.stderrs <- sqrt(jackknife.variance)
  return(jackknife.stderrs)
}
```



###Jackknife for Gamma Parameters Example

```{r, eval=FALSE}
input <- rgamma(1000, shape=0.1, scale=10)
gamma.est(input)
gamma.jackknife(input)
```



##Jackknife for linear regression coefficients

Write a function called `jackknife.lm` that takes arguments `df`, `formula` and `p` and returns jackknife standard error estimates on the coefficients of a linear regression model. 

```{r, eval=FALSE}
jackknife.lm <- function(df,formula,p) {
  n <- nrow(df)
  formula <- as.formula(formula)
  
  # Matrix of coefficients using omitted data
  theta.i <- matrix(nrow = n, ncol = p)
  for (row in 1:n) {
    theta.i[row,] <- lm(formula, data = df[-row,])$coefficients
  }
  
  # Calculates variance of estimates using omitted data
  variance <- apply(theta.i, 2, var)
  
  # Jackknife estimatation for variance and standard error
  jackknife.variance <- (n-1)^2/n * variance
  jackknife.stderr <- sqrt(jackknife.variance)
  return(jackknife.stderr)
}
```



###Jackknife for linear regression coefficients Example


```{r, eval=FALSE}
output <- 1.2 + 0.6*input +  rnorm(1000, 0, 2.1)
data <- data.frame(output,input)
my.lm <- lm(output~input, data=data)
# coefficients(my.lm)
# "Official" standard errors
sqrt(diag(vcov(my.lm)))
# Jackknife standard errors
jackknife.lm(df=data,formula="output~input",p=2)
```



##Refactoring the Jackknife


- Omitting one point or row is a common sub-task

- The general pattern:

```
figure out the size of the data
for each case
   omit that case
   repeat some estimation and get a vector of numbers
take variances across cases
scale up variances
take the square roots
```

- Refactor by extracting the common "omit one" operation

- Refactor by defining a general "jackknife" operation



###The Common Operation


- _Problem_: Omit one particular data point from a larger structure

- _Difficulty_: Do we need a comma in the index or not?

- _Solution_: Works for vectors, lists, 1D and 2D arrays, matrices, data frames:



## Goal: 


- Make the function select the correct dimensions
    - length for a 1d object
    - number of rows for 2d
- Write a function `omit.case` that omits a point given the data and returns the data minus that point. Make sure it can handle higher dimensions. 

```{r, eval=FALSE}
      omit.case <- function(the_data,omitted_point) {
        # Dim of vector is NULL ==> dim(1:4)
        # Dim of 1D array is vector with length 1 ==> dim(array(1:4), c(2,2)) 
        if (length(dim(the_data) == 1 || is.null(dim(the_data)))) {
          return(the_data[-omitted_point])
        } else {
          return(the_data[-omitted_point,])
        }
        
      }
```

- Write a function `omit_and_est` that takes the data with an omitted point and returns whatever function your estimator does. 


```{r, eval=FALSE}
jackknife <- function(estimator,the_data) {
  # Define n to be lenth or number of rows here
  if (length(dim(the_data)==1) || is.null(dim(the_data))) {
    n <- length(the_data)
  } else {
    n <- nrow(the_data)
  }
  
  omit_and_est <- function(omit) {
    estimator(omit.case(the_data, omit))
  }
  
  jackknife.ests <- matrix(sapply(1:n, omit_and_est), ncol=n)
  var.of.reestimates <- apply(jackknife.ests,1,var)
  jackknife.var <- ((n-1)^2/n)* var.of.reestimates
  jackknife.stderr <- sqrt(jackknife.var)
  return(jackknife.stderr)
}
```



---

###It works


```{r, eval=FALSE}
some_normals <- rnorm(100,mean=7,sd=5)
mean.jackknife <- function(a_vector) {
  n <- length(a_vector)
  theta.i <- vector(length = n)
  for (i in 1:n) {
    theta.i[i] <- mean(a_vector[-i])
  }
  variance <- var(theta.i)
  jackknife.variance <- (n-1)^2/n * variance
  jackknife.stderr <- sqrt(jackknife.variance)
  return(jackknife.stderr)
}
# jackknife(estimator=mean,the_data=some_normals)
all.equal(jackknife(estimator=mean,the_data=some_normals),
          mean.jackknife(some_normals))
```

---



```{r, eval=FALSE}
all.equal(jackknife(estimator=gamma.est,the_data=data$input),
          gamma.jackknife(data$input))
```

---


```{r, eval=FALSE}
all.equal(jackknife(estimator=gamma.est,the_data=data$input),
          gamma.jackknife(data$input), check.names=FALSE)
```


```{r, eval=FALSE}
est.coefs <- function(the_data) {
  return(lm(output~input,data=the_data)$coefficients)
}
# est.coefs(data)
all.equal(est.coefs(data), coefficients(my.lm))
```

---


```{r, eval=FALSE}
# jackknife(estimator=est.coefs,the_data=data)
all.equal(jackknife(estimator=est.coefs,the_data=data),
          jackknife.lm(df=data,formula="output~input",p=2))
```


##Further Refactoring of jackknife()


The code for `jackknife()` is still a bit clunky:
- Ugly `if-else` for finding `n`
- Bit at the end for scaling variances down to standard errors

- write a function that calculates the `n` needed for the above code:

    ```{r, eval=FALSE}
      data_size <- function(the_data) {
        if (length(dim(the_data)==1) || is.null(dim(the_data))) {
          return(length(the_data))
        } else {
          return(nrow(the_data))
        }
      }
    ```
- Write a funcrion that calculate the variance of all the estimates and returns the standard error
```{r, eval=FALSE}
scale_and_sqrt_vars <- function(jackknife.ests,n) {
  var.of.reestimates <- apply(jackknife.ests,1,var)
  jackknife.var <- ((n-1)^2/n)* var.of.reestimates
  jackknife.stderr <- sqrt(jackknife.var)
  return(jackknife.stderr)
}
```



###Now invoke those functions

```{r, eval=FALSE}
jackknife <- function(estimator,the_data) {
  n <- data_size(the_data)
  omit_and_est <- function(omit) {
    estimator(omit.case(the_data,omit))
  }
  jackknife.ests <- matrix(sapply(1:n, omit_and_est), ncol=n)
  return(scale_and_sqrt_vars(jackknife.ests,n))
}
```

